### 2021-11-16 src/main/r/MSDW_HPO_Phenotype_Analysis.Rmd
initial analysis

---
title: "LOINC2HPO_EPIC AND SCC"
output:
  html_document:
    df_print: paged
---

## GOAL
This is to LOINC2HPO analysis of both EPIC and SCC labs. 
```{r message=FALSE}
library(tidyverse)
library(glue)
library(readxl)
require(DBI)
require(RPostgres)
require(egg)
require(gridExtra)
source('.db_credential.R')
dsca <- DBI::dbConnect(RPostgres::Postgres(), host='localhost', user=user, dbname = dbname, port = 2345, password=password)
```

## ANALYSIS

### Prepare the lab tables

Run the following script to create the denormalized table for lab_scc and lab_epic. After this step, we should have three tables for SCC lab, EPIC lab, and combined. And each table should have LOINC and mapped HPO. 
```{r}
# sql_folder_path <- '/Users/xingminzhang/git/s4-pd-dmsdw/pd-dmsdw-loinc2hpo-app/src/main/resources/sql'
# sql_denormalized_scc_lab_2020july <- read_file(glue('{sql_folder_path}/loinc2hpo_SCC_MSDW.sql'))
# sql_denormalized_epic_lab_2020july <- read_file(glue('{sql_folder_path}/loinc2hpo_Epic_MSDW.sql'))
# sql_combine_Epic_SCC <- read_file(glue('{sql_folder_path}/loinc2hpo_combine_Epic_SCC.sql')) 
# 
# # make sure the lab table after LOINC2HPO transformation exists
# # if not, we can create it here; but the intended behavior is that we use the Java commandline app for this

```

```{sql connection=dsca, echo=FALSE}
drop table if exists denormalized_scc_lab_2020july;
```
```{sql connection=dsca}
CREATE TEMP TABLE denormalized_scc_lab_2020july as
WITH 
	lab_scc_2020q2 as (
		SELECT l.medical_record_number , l.encounter_key , l.calendar_date , l.time_of_day_key , l.clinical_result_numeric ,  trim(from l.unit_of_measure_numeric) as unit_of_measure_numeric , l.reference_range , l.procedure_key,
		case when l.abnormal_flag is null then 'N' else l.abnormal_flag end as abnormal_flag 
		FROM hai_az_prod.lab_scc_2020july l
		--where medical_record_number in (2966097, 3164683, 3869795)
	), 
	lab_scc_abnormal_flag_mapping AS (
		SELECT case when abf.abnormal_flag is NULL then 'N' else abf.abnormal_flag end as abnormal_flag, abf.mapto 
		FROM hai_az_prod.lab_scc_abnormal_flag_mapping abf
	)
SELECT lab.medical_record_number, lab.encounter_key , lab.calendar_date  as lab_date, dtod.clock_time_24_hour || ':00' as lab_time, lab.clinical_result_numeric , fdp.context_name, fdp.context_procedure_code AS local_test_code , loinc.loinc , lab.clinical_result_numeric AS lab_result_numeric_value, lab.unit_of_measure_numeric As unit_of_measure ,lab.reference_range AS normal_range , lab.abnormal_flag , abf.mapto AS interpretation, l2h.hpotermid , l2h.isnegated 
FROM lab_scc_2020q2 lab 
left JOIN hai_az_prod.fd_procedure fdp using (procedure_key)
left JOIN hai_az_prod.loinc_mapping loinc on fdp.context_procedure_code = loinc.code and (lab.unit_of_measure_numeric = loinc.unit or (lab.unit_of_measure_numeric is NULL and loinc.unit is NULL))
left JOIN lab_scc_abnormal_flag_mapping abf using (abnormal_flag)
left JOIN hai_az_prod.loinc2hpo l2h on loinc.loinc = l2h.loincid and abf.mapto = l2h.code
join prod_msdw.d_time_of_day dtod using (time_of_day_key);
```

```{sql connection=dsca}
DROP TABLE IF EXISTS denormalized_epic_lab_2020july;
```

```{sql connection=dsca}
create temp table denormalized_epic_lab_2020july as
with interpreted as (
	SELECT lab.mrn, lab.order_date , lab.lab_time , lab.test_code , lab.test_result_value , lab.unit_of_measurement , lab.reference_range ,
	case 
			when test_result_value_final :: REAL < range_low_final ::REAL then 'L' 
			when test_result_value_final :: REAL > range_hi_final :: REAL then 'H'
			else 'N' 
	end as interpretation
	FROM hai_az_test.lab_epic_2020july lab
	--where mrn in (3710567, 8014250, 3144637)
	)
SELECT lab.mrn AS medical_record_number, NULL AS encounter_key, lab.order_date as lab_date, lab.lab_time, 'EPIC' AS context_name, lab.test_code AS local_test_code , local2loinc.loinc , lab.test_result_value AS lab_result_numeric_value, lab.unit_of_measurement AS unit_of_measure , lab.reference_range AS normal_range , NULL AS abnormal_flag,  interpretation,
loinc2hpo.hpotermid , loinc2hpo.isnegated 
FROM interpreted lab
LEFT JOIN hai_az_prod.loinc_mapping local2loinc on lab.test_code :: VARCHAR = local2loinc.code and (lab.unit_of_measurement = local2loinc.unit OR (lab.unit_of_measurement is NULL and local2loinc.unit is NULL))
LEFT JOIN hai_az_prod.loinc2hpo loinc2hpo on local2loinc.loinc = loinc2hpo.loincid and interpretation = loinc2hpo.code 
where local2loinc."source" = 'EPIC'; 
```

Join the above tables together
```{sql connection=dsca}
CREATE VIEW denormalized_scc_and_epic_lab_2020july AS
WITH epic_denormalized AS (
  SELECT medical_record_number, lab_date, lab_time, context_name, local_test_code :: VARCHAR, loinc,  lab_result_numeric_value, unit_of_measure, normal_range, interpretation, hpotermid, isnegated
  FROM denormalized_epic_lab_2020july
), 
scc_denormalized AS (
  SELECT medical_record_number, lab_date, lab_time, context_name, local_test_code, loinc, lab_result_numeric_value, unit_of_measure, normal_range, interpretation, hpotermid, isnegated
  FROM denormalized_scc_lab_2020july
)
SELECT * FROM scc_denormalized 
UNION all 
SELECT * FROM epic_denormalized; 
```

Right now, we have three sets of denormalized labs, SCC, EPIC and their union.
```{sql connection=dsca}
SELECT * FROM denormalized_scc_lab_2020july limit 4;
```

```{sql connection=dsca}
SELECT * FROM denormalized_epic_lab_2020july limit 4;
```

```{r}
lab_tables <- c('denormalized_scc_lab_2020july', 'denormalized_epic_lab_2020july', 'denormalized_scc_and_epic_lab_2020july')
```

### Summary statistics before running any HPO inference

```{r}
abnormal_phenotypes_before_inference <- function(lab_table_name){
  results <- list()
  dbExecute(dsca, 'DROP TABLE IF EXISTS n_abnormal_before_infer_per_patient;')
  sql <- glue::glue('create temp table n_abnormal_before_infer_per_patient AS
        SELECT medical_record_number , count(*) as n_abnormal_before_infer, count(distinct hpotermid ) as n_uniq_abnormal_before_infer
        FROM {lab_table_name}
        where hpotermid is not null and not isnegated
        group by medical_record_number ;')
  dbExecute(dsca, sql)    
  # median of HPO-coded phenotypes per patient from abnormal lab findings
  median_abnormal_hpo <- dbGetQuery(dsca, 'SELECT median(n_abnormal_before_infer) FROM n_abnormal_before_infer_per_patient;')[1,1]
  results$n_abnormal_before_infer <- median_abnormal_hpo

  # median of unique HPO-coded phenotypes per patient from abnormal lab findings
  median_uniq_abnormal_before_infer <- dbGetQuery(dsca, 'SELECT median(n_uniq_abnormal_before_infer) FROM n_abnormal_before_infer_per_patient;')[1,1]
  results$n_uniq_abnormal_before_infer <- median_uniq_abnormal_before_infer

  n_abnormal_before_infer_per_patient <- dbGetQuery(dsca, 
        'SELECT n_abnormal_before_infer, count(*) as n_patient 
      FROM n_abnormal_before_infer_per_patient
      GROUP BY n_abnormal_before_infer
      order by n_abnormal_before_infer;')
  results$n_abnormal_before_infer_per_patient <- n_abnormal_before_infer_per_patient

  n_uniq_abnormal_before_infer_per_patient <- dbGetQuery(dsca, 
          'SELECT n_uniq_abnormal_before_infer, count(*) as n_patient 
            FROM n_abnormal_before_infer_per_patient
            GROUP BY n_uniq_abnormal_before_infer
            order by n_uniq_abnormal_before_infer;')
  results$n_uniq_abnormal_before_infer_per_patient <- n_uniq_abnormal_before_infer_per_patient

  return (results)
}

# the input is the third item in the above function
plot_patient_distribution_by_n_abnormal_phenotypes_before_infer <- function(df){
  n_abnormal_before_infer_per_patient <- df
  n_abnormal_before_infer_per_patient <- n_abnormal_before_infer_per_patient %>% mutate(n_abnormal_before_infer = as.integer(n_abnormal_before_infer), n_patient = as.integer(n_patient))

## roll up the ones > 100
  df_temp <- n_abnormal_before_infer_per_patient %>% mutate(n_abnormal_before_infer = ifelse(n_abnormal_before_infer > 100, 100, n_abnormal_before_infer)) %>% group_by(n_abnormal_before_infer) %>% summarise(n_patient = sum(n_patient))

  p <- ggplot(df_temp) + 
    geom_bar(aes(x=n_abnormal_before_infer, y = n_patient), stat='identity') +
    scale_x_continuous(limits = c(0, 105)) + 
    xlab("# HPO-coded phenotypes") + ylab("patient count") + 
    ggtitle("before HPO inference") +
    theme(axis.text = element_text(size = 8), title = element_text(size = 10))

  rm(df_temp)
  return (p)
}

plot_patient_distribution_by_n_uniq_abnormal_phenotypes_before_infer <- function(df){
  n_uniq_abnormal_before_infer_per_patient <- df
  n_uniq_abnormal_before_infer_per_patient <- n_uniq_abnormal_before_infer_per_patient %>% 
  mutate(n_uniq_abnormal_before_infer = as.integer(n_uniq_abnormal_before_infer), n_patient = as.integer(n_patient))

  p <- ggplot(n_uniq_abnormal_before_infer_per_patient) + 
    geom_bar(aes(x=n_uniq_abnormal_before_infer, y = n_patient), stat='identity') +
    scale_x_continuous(limits = c(0, 105)) +
    xlab("# HPO-coded phenotypes (unique)") + ylab("patient count") + 
    ggtitle("before HPO inference") +
    theme(axis.text = element_text(size = 8), title = element_text(size = 10))

  return (p)
}
```

```{r}
results_before_infer <- lapply(lab_tables, function(x) abnormal_phenotypes_before_inference(x))
```

Median of # of abnormal phenotypes before running HPO inference
```{r}
as.data.frame(list(lab_source = lab_tables, median_abnormal = sapply(results_before_infer, function(x) x$n_abnormal_before_infer), median_uniq_abnormal = sapply(results_before_infer, function(x) x$n_uniq_abnormal_before_infer)) )
```

What are most prevalent abnormal phenotypes? "Prevalent" is defined as unique number of patients having an abnormal phenotype. 
```{sql connection=dsca, max.output=-1}
SELECT l.hpotermid, hpo.label, l.n :: INTEGER
FROM 
      (SELECT hpotermid , count(distinct medical_record_number) as n
      FROM denormalized_scc_and_epic_lab_2020july
      WHERE hpotermid is not null and not isnegated
      group by hpotermid 
      order by n desc
      LIMIT 50) as l
LEFT JOIN hai_az_prod.hpo hpo on l.hpotermid = hpo.termid ;
```

### Summary statistics after running HPO inference

Count the number of HPO phenotypes per patient after inference. Note inference was only done for abnormal labs. It is tricky to infer based on normal findings.

```{r}
abnormal_phenotypes_after_inference <- function(lab_table_name){
  results <- list()
  dbExecute(dsca, 'DROP TABLE IF EXISTS n_abnormal_with_inferred_per_patient;')
  sql <- glue::glue('create temp table n_abnormal_with_inferred_per_patient AS
with abnormal_lab as 
	(SELECT * FROM {lab_table_name} where hpotermid is not null and not isnegated),
abnormal_lab_after_infer as 
	(SELECT l.medical_record_number, l.lab_date, l.local_test_code, l.loinc, l.hpotermid, l.isnegated, pair.ancestor , pair.distance 
	FROM abnormal_lab l 
	left join hai_az_prod.hpo_is_a_pairs pair on l.hpotermid = pair."current")
SELECT medical_record_number, count(*) as n_abnormal_with_inferred, count(DISTINCT ancestor) as n_uniq_abnormal_with_inferred
FROM abnormal_lab_after_infer 
group by medical_record_number;')
  dbExecute(dsca, sql)    
  # median of HPO-coded phenotypes per patient from abnormal lab findings
  median_abnormal_hpo <- dbGetQuery(dsca, 'SELECT median(n_abnormal_with_inferred) FROM n_abnormal_with_inferred_per_patient;')[1,1]
  results$n_abnormal_after_infer <- median_abnormal_hpo

  # median of unique HPO-coded phenotypes per patient from abnormal lab findings
  median_uniq_abnormal_after_infer <- dbGetQuery(dsca, 'SELECT median(n_uniq_abnormal_with_inferred) FROM n_abnormal_with_inferred_per_patient;')[1,1]
  results$n_uniq_abnormal_after_infer <- median_uniq_abnormal_after_infer

  n_abnormal_after_infer_per_patient <- dbGetQuery(dsca, 
        'SELECT n_abnormal_with_inferred, count(*) as n_patient 
        FROM n_abnormal_with_inferred_per_patient 
        group by n_abnormal_with_inferred 
        order by n_abnormal_with_inferred;')
  results$n_abnormal_before_infer_per_patient <- n_abnormal_after_infer_per_patient

  n_uniq_abnormal_after_infer_per_patient <- dbGetQuery(dsca, 
          'SELECT n_uniq_abnormal_with_inferred, count(*) as n_patient 
          FROM n_abnormal_with_inferred_per_patient 
          group by n_uniq_abnormal_with_inferred 
          order by n_uniq_abnormal_with_inferred;')
  results$n_uniq_abnormal_before_infer_per_patient <- n_uniq_abnormal_after_infer_per_patient

  dbExecute(dsca, 'DROP TABLE IF EXISTS n_abnormal_with_inferred_per_patient;')

  return (results)
}

# the input is the third item in the above function
plot_patient_distribution_by_n_abnormal_phenotypes_after_infer <- function(df){
  n_abnormal_with_inferred_per_patient <- df
  n_abnormal_with_inferred_per_patient <- n_abnormal_with_inferred_per_patient %>% mutate(n_abnormal_with_inferred = as.integer(n_abnormal_with_inferred), n_patient = as.integer(n_patient))
n_abnormal_with_inferred_per_patient %>% head()

  MAX_N_ABNORMAL_WITH_INFERRED <- max(n_abnormal_with_inferred_per_patient$n_abnormal_with_inferred)
  df_temp <- n_abnormal_with_inferred_per_patient %>% mutate(strata = cut(n_abnormal_with_inferred_per_patient$n_abnormal_with_inferred, breaks = c(seq(0,100, 10), seq(101, 1001, 100), seq(1002, MAX_N_ABNORMAL_WITH_INFERRED + 30000, 30000)))) %>% group_by(strata) %>% summarise(n_patient = sum(n_patient))

  p <- ggplot(df_temp) +
    geom_bar(aes(x = strata, y = n_patient), stat = 'identity') + theme(axis.text.x = element_text(angle = 30)) +
    xlab("# HPO-coded phenotypes") + ylab("patient count") + 
    ggtitle("after HPO inference") +
    theme(axis.text = element_text(size = 4), title = element_text(size = 10))

  rm(df_temp)
  return (p)
}

plot_patient_distribution_by_n_uniq_abnormal_phenotypes_after_infer <- function(df){
  n_uniq_abnormal_with_inferred_per_patient <- df
  n_uniq_abnormal_with_inferred_per_patient <- n_uniq_abnormal_with_inferred_per_patient %>% mutate(n_uniq_abnormal_with_inferred = as.integer(n_uniq_abnormal_with_inferred), n_patient = as.integer(n_patient))

  MAX_N_ABNORMAL_WITH_INFERRED <- max(n_uniq_abnormal_with_inferred_per_patient$n_uniq_abnormal_with_inferred)
  df_temp <- n_uniq_abnormal_with_inferred_per_patient %>% mutate(strata = cut(n_uniq_abnormal_with_inferred, breaks = c(seq(0,100, 10), seq(101, 1001, 100), seq(1002, MAX_N_ABNORMAL_WITH_INFERRED + 30000, 30000)))) %>% group_by(strata) %>% summarise(n_patient = sum(n_patient))

  p <- ggplot(df_temp) +
    geom_bar(aes(x = strata, y = n_patient), stat = 'identity') + theme(axis.text.x = element_text(angle = 45)) +
    xlab("# HPO-coded phenotypes (unique)") + ylab("patient count") + 
    ggtitle("after HPO inference") +
    theme(axis.text = element_text(size = 4), title = element_text(size = 10))
  rm(df_temp)

  return (p)
}
```

```{r}
results_after_infer <- lapply(lab_tables, abnormal_phenotypes_after_inference)
```

```{r}
as.data.frame(list(lab_source = lab_tables, median_abnormal = sapply(results_after_infer, function(x) x$n_abnormal_after_infer), median_uniq_abnormal = sapply(results_after_infer, function(x) x$n_uniq_abnormal_after_infer)) )
```

### Show patient distribution plots before and after inference on one plot

```{r}
# input is the list item in results_before_infer, and results_after_infer
plot_grid <- function(result_before_infer, result_after_infer){
  # patient distribution ~ abnormal phenotypes before inference
  df_abnormal_before <- result_before_infer[[3]]
  p1 <- plot_patient_distribution_by_n_abnormal_phenotypes_before_infer(df_abnormal_before)
  # patient distribution ~ uniq abnormal phenotypes before inference
  df_uniq_abnormal_before <- result_before_infer[[4]]
  p2 <- plot_patient_distribution_by_n_uniq_abnormal_phenotypes_before_infer(df_uniq_abnormal_before)
  # patient distribution ~ abnormal phenotypes after inference
  df_abnormal_after <- result_after_infer[[3]]
  p3 <- plot_patient_distribution_by_n_abnormal_phenotypes_after_infer(df_abnormal_after)
  # patient distribution ~ uniq abnormal phenotypes after inference
  df_uniq_abnormal_after <- result_after_infer[[4]]
  p4 <- plot_patient_distribution_by_n_uniq_abnormal_phenotypes_after_infer(df_uniq_abnormal_after)

  width = unit(8, "cm")
  height = unit(3, "cm")
  p <- gridExtra::grid.arrange(
    set_panel_size(p1, width = width, height = height),
    set_panel_size(p2, width = width, height = height),
    set_panel_size(p3, width = width, height = height),
    set_panel_size(p4, width = width, height = height),
    ncol=2)

  return (p)
}
```

Patient distribution by looking at SCC labs alone
```{r}
# Lab_SCC
plot_grid(results_before_infer[[1]], results_after_infer[[1]])
```

Patient distribution by looking at EPIC labs alone
```{r}
# Lab_EPIC
plot_grid(results_before_infer[[2]], results_after_infer[[2]])
```

Patient distribution by looking at all labs combined
```{r}
# Lab_SCC_and_EPIC
plot_grid(results_before_infer[[3]], results_after_infer[[3]])
```

We can also look at the most prevalent phenotypes after inference. It is not fair to compare all labs because parent terms are always going to be more prevalent. So we just look at level 1 terms (terms right below "Phenotypic Abnormality").
Look at patients at level = 1 (right below phenotypic abnormality)
```{sql connection=dsca, max.output=-1}
with abnormal_lab as 
	(SELECT * FROM denormalized_scc_and_epic_lab_2020july where hpotermid is not null and not isnegated),
abnormal_lab_after_infer as 
	(SELECT l.medical_record_number, l.lab_date, l.local_test_code, l.loinc, l.hpotermid, l.isnegated, pair.ancestor , pair.distance , hpo.distancetoroot, hpo.label
	FROM abnormal_lab l 
	left join hai_az_prod.hpo_is_a_pairs pair on l.hpotermid = pair."current"
	left join hai_az_prod.hpo hpo on pair.ancestor = hpo.termid)
SELECT ancestor, label, count(distinct medical_record_number) :: INTEGER as n_patient FROM abnormal_lab_after_infer where distancetoroot = 1 group by ancestor, label
ORDER BY n_patient DESC;
```

```{sql connection=dsca}
DROP VIEW IF EXISTS denormalized_scc_and_epic_lab_2020july;
```

### Analyze the tested phenotypes

Whether a lab test is ordered or not can sometimes be very informative. In this section, we will analyze the most tested phenotypes.

Some phenotypes can be directly tested, such as "Increased LDL cholesterol concentration". Other phenotypes can be inferred from directly testable phenotypes, such as "Abnormal Leukocyte count", which are inferred from specific tests on different blood cell counts. We collect the count of phenotypes directly tested, or including indirectly tested ones as well.

```{r}
CACHE_DIR = 'cache'
if (!dir.exists(CACHE_DIR)){
  dir.create(CACHE_DIR)
}

get_tested_phenotype_rank_direct <- function(force_rerun=FALSE, cache_dir=CACHE_DIR){
  df_path = glue('{cache_dir}/tested_phenotype_rank_direct.csv')

  if (force_rerun | !file.exists(df_path)){
    df <- dbGetQuery(dsca, "
      with testable_phenotypes_direct as (
    	select lab.medical_record_number , lab.context_name , lab.lab_date , lab.lab_time , lab.loinc , testable.termid , testable.is_tested_directly 
    	from hai_az_prod.scc_and_epic_lab_after_loinc2hpo_2020july lab
    	join hai_az_prod.loinc_testable_phenotypes testable on lab.loinc = testable.loincid 
    	where testable.is_tested_directly 
    ),
    testable_phenotype_direct_freq as (
    	select termid, count(*) as n_record, count(distinct medical_record_number) as n_patient
    	from testable_phenotypes_direct
    	group by termid 
    )
    select testable.*, hpo.label
    from testable_phenotype_direct_freq testable
    join hai_az_prod.hpo hpo on testable.termid = hpo.termid
    order by n_record desc;")

    write.csv(df, file = df_path, row.names=FALSE)
  }

  df <- read.csv(df_path)

  return (df)
}

get_tested_phenotype_rank_all <- function(force_rerun=FALSE, cache_dir=CACHE_DIR){
  df_path = glue('{cache_dir}/tested_phenotype_rank_all.csv')

  if (force_rerun | !file.exists(df_path)){
    df <- dbGetQuery(dsca, "
               with testable_phenotypes_direct as (
              	select lab.medical_record_number , lab.context_name , lab.lab_date , lab.lab_time , lab.loinc , testable.termid , testable.is_tested_directly 
              	from hai_az_prod.scc_and_epic_lab_after_loinc2hpo_2020july lab
              	join hai_az_prod.loinc_testable_phenotypes testable on lab.loinc = testable.loincid 
              	--where testable.is_tested_directly 
              ),
              testable_phenotype_direct_freq as (
              	select termid, count(*) as n_record, count(distinct medical_record_number) as n_patient
              	from testable_phenotypes_direct
              	group by termid 
              )
              select testable.*, hpo.label, hpo.distancetoroot 
              from testable_phenotype_direct_freq testable
              join hai_az_prod.hpo hpo on testable.termid = hpo.termid
              order by n_record desc;
              ")

    write.csv(df, file = df_path, row.names=FALSE)
  }

  df <- read.csv(df_path)

  return (df)
}

tested_phenotype_rank_direct <- get_tested_phenotype_rank_direct()
tested_phenotype_rank_all <- get_tested_phenotype_rank_all()
```

For the directly tested phenotypes, we can observe what are the most tested abnormalities in one healthcare system. 
```{r}
tested_phenotype_rank_direct %>% head(10) %>% formattable::formattable()
```

For the directly/indirectly tested phenotypes, we can determine at the higher level, what are the most tested abnormalities. 
```{r max.output=-1}
tested_phenotype_rank_all %>% filter(distancetoroot == 1) %>% arrange(-n_record) %>% select(-distancetoroot)
```

For both cases, it would be interesting to learn about the changes of the lab phenotypes over time, and among different races.

```{r}
get_tested_phenotype_rank_direct_by_race <- function(force_rerun=FALSE, cache_dir=CACHE_DIR){
  df_path = glue('{cache_dir}/tested_phenotype_rank_direct_by_race.csv')

  if (force_rerun | !file.exists(df_path)){
    df <- dbGetQuery(dsca, "
              with testable_phenotypes_direct as (
            	select lab.medical_record_number , person.race, lab.context_name , lab.lab_date , lab.lab_time , lab.loinc , testable.termid , testable.is_tested_directly 
            	from hai_az_prod.scc_and_epic_lab_after_loinc2hpo_2020july lab
            	join hai_az_prod.loinc_testable_phenotypes testable on lab.loinc = testable.loincid 
            	join hai_az_prod.person_msdw_2020july person using (medical_record_number)
            	where testable.is_tested_directly 
            ),
            testable_phenotype_direct_freq as (
            	select termid, race, count(*) as n_record, count(distinct medical_record_number) as n_patient
            	from testable_phenotypes_direct
            	group by termid, race
            )
            select testable.*, hpo.label
            from testable_phenotype_direct_freq testable
            join hai_az_prod.hpo hpo on testable.termid = hpo.termid
            order by termid, race;")

    write.csv(df, file = df_path, row.names=FALSE)
  }

  df <- read.csv(df_path)

  return (df)
}

get_tested_phenotype_rank_all_by_race <- function(force_rerun=FALSE, cache_dir=CACHE_DIR){
  df_path = glue('{cache_dir}/tested_phenotype_rank_all_by_race.csv')

  if (force_rerun | !file.exists(df_path)){
    df <-  dbGetQuery(dsca, "
              with testable_phenotypes_direct as (
              	select lab.medical_record_number , person.race, lab.context_name , lab.lab_date , lab.lab_time , lab.loinc , testable.termid , testable.is_tested_directly 
              	from hai_az_prod.scc_and_epic_lab_after_loinc2hpo_2020july lab
              	join hai_az_prod.loinc_testable_phenotypes testable on lab.loinc = testable.loincid 
              	join hai_az_prod.person_msdw_2020july person using (medical_record_number)
              	--where testable.is_tested_directly 
              ),
              testable_phenotype_direct_freq as (
              	select termid, race, count(*) as n_record, count(distinct medical_record_number) as n_patient
              	from testable_phenotypes_direct
              	group by termid, race
              )
              select testable.*, hpo.label, hpo.distancetoroot
              from testable_phenotype_direct_freq testable
              join hai_az_prod.hpo hpo on testable.termid = hpo.termid
              order by termid, race;")

    write.csv(df, file = df_path, row.names=FALSE)
  }

  df <- read.csv(df_path)

  return (df)
}

tested_phenotype_rank_direct_by_race <- get_tested_phenotype_rank_direct_by_race()
tested_phenotype_rank_all_by_race <- get_tested_phenotype_rank_all_by_race()
```

For each directly measured phenotype, we can determine whether race and the tests are correlated by some form of Chi-squared teste.
```{r}
tested_phenotype_rank_direct_by_race %>% filter(race == 'African American') %>% arrange(-n_record) %>% head(10) %>% formattable::formattable()
tested_phenotype_rank_direct_by_race %>% filter(race == 'Asian') %>% arrange(-n_record) %>% head(10) %>% formattable::formattable()
tested_phenotype_rank_direct_by_race %>% filter(race == 'White') %>% arrange(-n_record) %>% head(10) %>% formattable::formattable()
```

```{r}
tested_phenotype_rank_all_by_race %>% filter(race == 'African American' & distancetoroot == 1) %>% arrange(-n_record) %>% head(10) %>% formattable::formattable()
tested_phenotype_rank_all_by_race %>% filter(race == 'Asian' & distancetoroot == 1) %>% arrange(-n_record) %>% head(10) %>% formattable::formattable()
tested_phenotype_rank_all_by_race %>% filter(race == 'White' & distancetoroot == 1) %>% arrange(-n_record) %>% head(10) %>% formattable::formattable()
```

Plot the anual change of tested phenotypes.

```{r}
get_tested_phenotype_rank_all_by_year <- function(force_rerun=FALSE, cache_dir=CACHE_DIR){
  df_path = glue('{cache_dir}/tested_phenotype_rank_all_by_year.csv')

  if (force_rerun | !file.exists(df_path)){
    df <-  dbGetQuery(dsca, "
              with testable_phenotypes_direct as (
              	select lab.medical_record_number , person.race, lab.context_name , lab.lab_date , lab.lab_time , date_part(year, lab_date) as lab_year, lab.loinc , testable.termid , testable.is_tested_directly 
              	from hai_az_prod.scc_and_epic_lab_after_loinc2hpo_2020july lab
              	join hai_az_prod.loinc_testable_phenotypes testable on lab.loinc = testable.loincid 
              	join hai_az_prod.person_msdw_2020july person using (medical_record_number)
              	--where testable.is_tested_directly 
              ),
              testable_phenotype_direct_freq as (
              	select termid, lab_year, count(*) as n_record, count(distinct medical_record_number) as n_patient
              	from testable_phenotypes_direct
              	group by termid, lab_year
              )
              select testable.*, hpo.label, hpo.distancetoroot 
              from testable_phenotype_direct_freq testable
              join hai_az_prod.hpo hpo on testable.termid = hpo.termid
              order by termid, lab_year;
              ")

    write.csv(df, file = df_path, row.names=FALSE)
  }

  df <- read.csv(df_path)

  return (df)
}

tested_phenotype_rank_all_by_year <- get_tested_phenotype_rank_all_by_year()

plot_phenotypes_of_interest_by_year <- function(terms_of_interest, df = tested_phenotype_rank_all_by_year){
  hpo_terms_selected <- df %>% filter(termid %in% terms_of_interest) %>% select(termid, label) %>% distinct()

  p <- df %>% filter(termid %in% terms_of_interest) %>% 
    filter(lab_year <= 2019 & lab_year > 2000) %>%
    ggplot(aes(x = lab_year, y = n_patient, group = termid, color = termid)) + 
    geom_point() +
    geom_line() +
    scale_color_discrete(breaks = hpo_terms_selected$termid, labels = str_c(hpo_terms_selected$termid, hpo_terms_selected$label, sep = " "))

  return (p)
}
```

```{r}
hpo_level_1_terms <- tested_phenotype_rank_all_by_year %>% filter(distancetoroot == 1) %>% select(termid, label) %>% distinct()
plot_phenotypes_of_interest_by_year(terms_of_interest = hpo_level_1_terms$termid)
```

Some trends are pretty interesting, such as the spike of Abnormal cellular phenotype. The term has three children, but only one is tested: HP:0011017 Abnormal cellular physiology

```{r}
children_abnormal_cellular_physiology <- c('HP:0003220', 'HP:0011018', 'HP:0025463', 'HP:0011019', 'HP:0003254', 'HP:4000056', 'HP:0011133', 'HP:0004371', 'HP:0031377', 'HP:0031409', 'HP:0012103', 'HP:0032576', 'HP:0003204', 'HP:0004356', 'HP:0032571', 'HP:0002916', 'HP:0003224', 'HP:0003575', 'HP:0003358')
plot_phenotypes_of_interest_by_year(terms_of_interest = children_abnormal_cellular_physiology)
```

Again, the above plot shows that it is driven by one HP:0031409 Abnormal lymphocyte physiology. 
```{r}
children_of_interest <- c('HP:0011840', 'HP:0030886', 'HP:0012177', 'HP:0005372', 'HP:0031378')
plot_phenotypes_of_interest_by_year(terms_of_interest = children_of_interest)
```
```{r}
children_of_interest <- c('HP:0002847', 'HP:0002959', 'HP:0005384', 'HP:0041076', 'HP:0010701')
plot_phenotypes_of_interest_by_year(terms_of_interest = children_of_interest)
```

```{r}
children_of_interest <- c('HP:0010702', 'HP:0032336', 'HP:0410241', 'HP:0410240', 'HP:0410243', 'HP:0410242', 'HP:0410244', 'HP:0004313')
plot_phenotypes_of_interest_by_year(terms_of_interest = children_of_interest)
```

```{r}
children_of_interest <- c('HP:0410367', 'HP:0031047', 'HP:0003237', 'HP:0003212', 'HP:0003261', 'HP:0410369', 'HP:0410371', 'HP:0410246', 'HP:0003496')
plot_phenotypes_of_interest_by_year(terms_of_interest = children_of_interest)
```