### 2021-12-01 src/main/r/numeric_vs_hpo_for_acute_kidney_failure.Rmd
compare raw and hpo of kidney labs for acute kidney failure

---
title: "R Notebook"
output: html_notebook
---

## Goal
Compare the performance of raw lab values vs HPO interpretations for acute kidney failure.

The analysis process is very similar to the liver function analysis.

Algorithm:
- Define a cohort of patients coded with acute kidney failure as the primay diagnosis of a visit;
- Get all the lab tests with selected LOINC code; 
- For each patient, the labs within X days (e.g. 7 days) before the diagnosis date is considered to be associated with + outcome; and the labs further outside of the window is considered to be associated with - outcome; 
- For each LOINC code, we will prepare a dataframe, separate into training and test set, train a logistic regression classifier using either raw numerical value or HPO value (observed or not observed), build the ROC curve with the test set, and finally compare the ROC curve to determine whether they are statistically different.

## Analysis

### retrieve data
```{r message=FALSE}
library(tidyverse)
library(glue)
library(lubridate)
library(readxl)
require(DBI)
require(RPostgres)
require(egg)
require(gridExtra)
source('.db_credential.R')
source('functions.R')
dsca <- DBI::dbConnect(RPostgres::Postgres(), host='localhost', user=user, dbname = dbname, port = 2345, password=password)

# cache directory to read from or write to
CACHE_DIR <- './cache'
if (!dir.exists(CACHE_DIR)){
  dir.create(CACHE_DIR)
}
```

Create a cohort for patients with abnormal liver function diagnoses.
```{r}
sql <- read_file('../sql/acute_kidney_failure_cohort.sql')
print(sql)
n<- dbExecute(dsca, sql)
```

Retrieve liver function diagnoses information for the cohort.
```{r}
sql <- "SELECT * FROM cohort"
df_cohort <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = 'acute_kidney_failure_diagnoses.csv', db_access_callback = dbGetQuery, conn = dsca, statement = sql, force_rerun = FALSE)

df_cohort %>% nrow()
```

Retrieve liver function lab tests for the cohort. 
```{r}
sql <- read_file('../sql/get_labs_for_acute_kidney_failure.sql')
df_labs <- get_df_from_db_or_cache(cache_dir = CACHE_DIR, file_name = "labs_acute_kidney_failure.csv", db_access_callback = dbGetQuery, conn=dsca, statement = sql, force_rerun = FALSE)
df_labs %>% nrow()
```

The first thing is to check how do the raw values of lab test look like before diagnoses of abnormal liver function. We want to choose a time window during which the liver function labs are highly correlated with the diagnosis, and another time window well, which is before the first window and corresponds to a *normal* liver function labs. The plot below shows that roughly we can choose -180 as the time cutoff: within this time cutoff, i.e. (-180, 0) days, the latest lab test tends to be abnormal and correspond to an abnormal liver function diagnosis; outside of this time cutoff, i.e. 180 days before a known liver abnormal function diagnosis, the lab test tends to be normal.  
```{r}
plot_lab_raw_values_vs_rel_date <- function(lab_loinc, ylim=NULL){
  df <- df_labs %>% 
    filter(loinc == lab_loinc) %>% 
    filter(rel_date <= 0 & rel_date > -730) %>% 
    mutate(lab_result_numeric_value = as.numeric(lab_result_numeric_value))

  p <- ggplot(df, aes(x = rel_date, y = lab_result_numeric_value)) +
    geom_point(size = 0.1, alpha=0.1) +
    xlab('lab date (relative to diagnosis)')

  if (!is.null(ylim)){
    p <- p + coord_cartesian(ylim = ylim)
  }

  p <- p + theme_bw() + 
    theme(panel.grid = element_blank(),
          axis.text = element_text(size = 5),
          axis.title = element_text(size = 8), 
          strip.text.x = element_text(margin = margin(0.3, 0, 0.3, 0), size = 5),
          plot.title = element_text(size = 8))

  return (p)
}

# setClass('LiverLab', representation(lab_loinc = 'character', lab_name = 'character', ylim_min = "numeric", ylim_max = "numeric"))
# liver_lab_ALT <- new('LiverLab', lab_loinc = '1742-6', lab_name = 'Alanine aminotransferase', ylim_min = 0, ylim_max = 500)
# liver_lab_GGT <- new('LiverLab', lab_loinc = '2324-2', lab_name = 'Gamma glutamyl transferase', ylim_min = 0, ylim_max = 1000)
# liver_lab_AST <- new('LiverLab', lab_loinc = '1920-8', lab_name = 'Aspartate aminotransferase', ylim_min = 0, ylim_max = 500)
# liver_lab_ALP <- new('LiverLab', lab_loinc = '6768-6', lab_name = 'Alkaline phosphatase', ylim_min = 0, ylim_max = 1000)
# liver_lab_total_protein <- new('LiverLab', lab_loinc = '2885-2', lab_name = 'Total protein', ylim_min = 0, ylim_max = 15)
# liver_lab_serum_albumin <- new('LiverLab', lab_loinc = '1751-7', lab_name = 'Serum albumin', ylim_min = 0, ylim_max = 20)
# liver_lab_PT <- new('LiverLab', lab_loinc = '5902-2', lab_name = 'Prothrombin time', ylim_min = 0, ylim_max = 50)
# liver_lab_total_bulirubin <- new('LiverLab', lab_loinc = '1975-2', lab_name = 'Total bilirubin', ylim_min = 0, ylim_max = 15)
# liver_lab_direct_bulirubin <- new('LiverLab', lab_loinc = '1968-7', lab_name = 'Direct bilirubin', ylim_min = 0, ylim_max = 15)
# 
# liver_labs <- c(liver_lab_ALT, 
#                 liver_lab_GGT, 
#                 liver_lab_AST, 
#                 liver_lab_ALP, 
#                 liver_lab_total_protein, 
#                 liver_lab_serum_albumin, 
#                 liver_lab_total_bulirubin, 
#                 liver_lab_direct_bulirubin)
# 
# plots <- map(liver_labs, function(x) plot_lab_raw_values_vs_rel_date(lab_loinc = x@lab_loinc, ylim = c(x@ylim_min, x@ylim_max)))
# p <- arrangeGrob(sapply(plots, function(x) x[[1]]), ncol = 2)

p1 <- plot_lab_raw_values_vs_rel_date(lab_loinc = '2160-0', ylim = c(0, 20)) + ggtitle('Serum creatinine')

p2 <- plot_lab_raw_values_vs_rel_date(lab_loinc = '3094-0', ylim = c(0, 150)) + ggtitle('Urea nitrogen in serum')

p3 <- plot_lab_raw_values_vs_rel_date(lab_loinc = '6299-2', ylim = c(0, 200)) + ggtitle('Urea nitrogen in blood')

p4 <- plot_lab_raw_values_vs_rel_date(lab_loinc = '2823-3', ylim = c(0, 10)) + ggtitle('Potassium in serum')

# seems the test strip only reports positive or negative outcomes
p5 <- plot_lab_raw_values_vs_rel_date(lab_loinc = '5804-0', ylim = c(0, 500)) + ggtitle('protein in urine')

# very sparse data
p6 <- plot_lab_raw_values_vs_rel_date(lab_loinc = '2889-4', ylim = c(0, 200)) + ggtitle('Protein in 24h urine')

# very very sparse data
p7 <- plot_lab_raw_values_vs_rel_date(lab_loinc = '2888-6', ylim = c(0, 500)) + ggtitle('Protein in urine')

p <- gridExtra::arrangeGrob(p1, p2, p3, p4, p5, p6, p7, ncol = 3)

ggsave(filename = 'cache/kidney_labs_raw_data_distribution_by_rel_date.png', plot = p, width = 6, height = 6)
```

![liver function labs](./cache/kidney_labs_raw_data_distribution_by_rel_date.png)

Use -180 as the time cutoff and separate the patient journey into two sections: 0 - 180 days before the first diagnosis of abnormal liver function, and 180 days before the first diagnosis of liver abnormal functions. Within each section, we will pick the latest lab measurement for each LOINC lab test (features) and assign either + (within 180 days) or - outcome (outside of 180 days). 
```{r}
# prepare data entries for + and - outcomes
# each lab result is associated with an outcome
# briefly, the following snippet uses the following algorithm:
# calculate the relative date to the first abnormal liver function diagnosis: if before diagnosis, - days otherwise + days
# then cut the relative days into a few windows (-Inf, -365), (-365, -30), (-30, -7), (-7, 0) and (0, Inf)
# for each window, we only use the closest one to diagnosis date as a representative, although we can use other methods such as the mean
# at last, only the (-7, 0) window has + outcome, meaning having abnormal liver function; all other windows have normal liver function

df_labs_and_outcomes <- df_labs %>% filter(loinc %in% c('2160-0', '3094-0', '6299-2', '2823-3')) %>%
  as_tibble() %>%
  select(-min_diag_date, -rel_date) %>% ## these two columns can be removed from the sql query
  ##filter(medical_record_number %in% c('2715515', '3087585')) %>% ## test with a few patient
  left_join(df_cohort, by = 'medical_record_number') %>%
  mutate(lab_date = ymd(lab_date), min_diag_date = ymd(min_diag_date)) %>%
  mutate(lab_date_rel_to_diag = as.duration(interval(min_diag_date, lab_date)) / ddays()) %>%
  mutate(window_num = cut(lab_date_rel_to_diag, c(-Inf, -180, 0, Inf))) %>%
##select(medical_record_number, loinc, lab_date, min_diag_date, lab_date_rel_to_diag, window_num)
  filter(window_num %in% c('(-Inf,-180]', '(-180,0]')) %>%
  arrange(medical_record_number, loinc, -lab_date_rel_to_diag) %>%
  group_by(medical_record_number, loinc, window_num) %>%
  mutate(sequence = row_number()) %>%
  #select(medical_record_number, loinc, lab_date, min_diag_date, lab_date_rel_to_diag, window_num, sequence) %>%
  ungroup() %>%
  filter(sequence == 1) %>%
  mutate(outcome = if_else(window_num == '(-180,0]', 'abnormal', 'normal'))

# the following shows that we have pretty balanced classes for each lab test
df_labs_and_outcomes %>% group_by(loinc, outcome) %>% summarise(n = n())
```

Define a list of functions to run statistical comparisons.

```{r}
require(caret)
require(broom)
require(pROC)

# preprocess: convert the HPO column (and the isnegated column) into three separate features:
# is_hpoterm_observed_for_H, is_hpoterm_observed_for_L and is_hpoterm_observed_for_N
# pass in the dataframe of lab tests of one LOINC, return a new dataframe with the three new features
create_hpo_features <- function(df_one_loinc){
  df_hpo_features <- df_one_loinc %>%
    mutate(dummy_value = 1) %>%
    pivot_wider(names_from = interpretation, names_prefix = 'is_hpoterm_observed_for_', values_from = dummy_value, values_fill = 0) %>%
    select(lab_result_numeric_value, starts_with('is_hpoterm_observed_for_'), outcome)

  df_hpo_features <- df_hpo_features %>%
    mutate(outcome = as.factor(outcome)) %>%
    mutate(outcome = relevel(outcome, 'normal')) %>%
    mutate(lab_result_numeric_value = as.numeric(lab_result_numeric_value))

  return (df_hpo_features)
}

train_test_split <- function(df, p, ...){
  trainIndex = createDataPartition(df$outcome, p = p, ...)
  df_train <- df[trainIndex, ]
  df_test <- df[-trainIndex, ]
  return (list(df_train=df_train, df_test=df_test))
}

lr_for_hpo_H <- function(df_train, df_test){
  model_continuous <- glm(formula = 'outcome ~ lab_result_numeric_value', data = df_train, family = binomial(link = 'logit'))
  model_hpo <- glm(formula = 'outcome ~ is_hpoterm_observed', data = df_train %>% rename(is_hpoterm_observed = is_hpoterm_observed_for_H), family = binomial(link = 'logit'))
  prediction_continuous <- predict(model_continuous, df_test, type = 'response')
  prediction_hpo <- predict(model_hpo, df_test %>% rename(is_hpoterm_observed = is_hpoterm_observed_for_H), type = 'response')
  roc_continuous <- roc(df_test$outcome, prediction_continuous)
  roc_hpo <- roc(df_test$outcome, prediction_hpo)
  return (list(model_continuous, model_hpo, prediction_continuous, prediction_hpo, roc_continuous, roc_hpo))
}

lr_for_hpo_L <- function(df_train, df_test){
  model_continuous <- glm(formula = 'outcome ~ lab_result_numeric_value', data = df_train, family = binomial(link = 'logit'))
  model_hpo <- glm(formula = 'outcome ~ is_hpoterm_observed', data = df_train %>% rename(is_hpoterm_observed = is_hpoterm_observed_for_L), family = binomial(link = 'logit'))
  prediction_continuous <- predict(model_continuous, df_test, type = 'response')
  prediction_hpo <- predict(model_hpo, df_test %>% rename(is_hpoterm_observed = is_hpoterm_observed_for_L), type = 'response')
  roc_continuous <- roc(df_test$outcome, prediction_continuous)
  roc_hpo <- roc(df_test$outcome, prediction_hpo)
  return (list(model_continuous, model_hpo, prediction_continuous, prediction_hpo, roc_continuous, roc_hpo))
}

compare_rocs <- function(roc1, roc2){
  results <- broom::tidy(roc.test(roc1, roc2, method = "bootstrap"))
  return (results)
}

pipeline_compare_for_outcome_prediction <- function(df_one_loinc) {

  df_hpo_features <- create_hpo_features(df_one_loinc = df_one_loinc)
  df_train_test <- train_test_split(df_hpo_features, p = 0.7, list=FALSE)

  df_train <- df_train_test$df_train
  df_test <- df_train_test$df_test

  # check if we have H for analysis, and/or L
  valid_lab_outcomes <- df_one_loinc %>% filter(!is.na(hpotermid) & !isnegated) %>% select(interpretation, hpotermid) %>% distinct()

  roc_comparisons <- list()
  if ('H' %in% valid_lab_outcomes$interpretation){
    lr_H <- lr_for_hpo_H(df_train = df_train, df_test = df_test)
    roc_comparison <- compare_rocs(lr_H[[5]], lr_H[[6]])
    roc_comparison <- bind_cols(roc_comparison, valid_lab_outcomes %>% filter(interpretation == 'H'))
    roc_comparisons$H <- roc_comparison
  }

  if ('L' %in% valid_lab_outcomes$interpretation){
    lr_L <- lr_for_hpo_L(df_train = df_train, df_test = df_test)
    roc_comparison <- compare_rocs(lr_L[[5]], lr_L[[6]])
    roc_comparison <- bind_cols(roc_comparison, valid_lab_outcomes %>% filter(interpretation == 'L'))
    roc_comparisons$L <- roc_comparison
  }

  roc_comparisons_combined <- bind_rows(roc_comparisons)

  roc_comparisons_combined
}

example_loinc = '2160-0'
df_one_loinc <- df_labs_and_outcomes %>% filter(loinc == example_loinc)
pipeline_compare_for_outcome_prediction(df_one_loinc = df_one_loinc)
```

```{r}
require(furrr)
plan(multisession, workers = 4)

kidney_analysis <- df_labs_and_outcomes %>%
  group_by(loinc) %>%
  nest() %>%
  mutate(result = future_map(data, pipeline_compare_for_outcome_prediction)) %>%
  select(-data) %>%
  unnest(result)

kidney_analysis
```

```{r}
df_one_loinc <- df_labs_and_outcomes %>% filter(loinc == '1742-6')

df_hpo_features <- create_hpo_features(df_one_loinc = df_one_loinc)

df_train_test <- train_test_split(df_hpo_features, p = 0.7, list=FALSE)

df_train <- df_train_test$df_train
df_test <- df_train_test$df_test

lr_H <- lr_for_hpo_H(df_train = df_train, df_test = df_test)
roc_comparison <- compare_rocs(lr_H[[5]], lr_H[[6]])
roc_comparison <- bind_cols(roc_comparison, valid_lab_outcomes %>% filter(interpretation == 'H'))
roc_comparisons$H <- roc_comparison

plot(lr_H[[5]])
plot(lr_H[[6]])

roc.test(lr_H[[5]], lr_H[[6]], method = "bootstrap")
```