### 2022-04-21 code/r/classifier_severe_asthma.Rmd
reorg project structure

---
title: "classifier_severe_asthma"
author: "Aaron Zhang"
date: "3/30/2022"
output: html_document
---

The goal of this document is to build ML classifiers to predict progression into severe asthma using biomarkers idenfied from lab phenotype screening.

```{r message=FALSE}
library(tidyverse)
library(glue)
```

## Load data
```{r}
CACHE_DIR <- '../cache'

confounders_outcome <- read_csv(glue("{CACHE_DIR}/confounders_outcome.csv"))
phenotype_features <- read_csv(glue("{CACHE_DIR}/phenotype_features.csv"))
```

## Prepare feature matrix

We will use the biomarker in the past year, and the count in the past three
```{r}
biomarkers_identified <- c("HP:0001880", "HP:0001899", "HP:0012312", 
                           "HP:0032309", "HP:0011897", "HP:0012416")

# biomarkers from Year (-1, 0] only
biomarker_feature_matrix <- phenotype_features %>% filter(termid %in% biomarkers_identified) %>% 
  mutate(termid = make.names(termid)) %>% 
  rename(tested = tested_count) %>% 
  rename(observed = observed_count) %>% 
  select(medical_record_number, lab_date_rel_bracket, termid, tested, observed) %>% 
  #filter(lab_date_rel_bracket == '(-1,0]') %>% 
  pivot_wider(id_cols = medical_record_number, names_from = c(lab_date_rel_bracket, termid), values_from = c(tested, observed))

Xy <- confounders_outcome %>% 
  left_join(biomarker_feature_matrix, by = 'medical_record_number') %>% 
  mutate(across(starts_with('tested'), ~ replace_na(.x, 0))) %>% 
  mutate(across(starts_with('observed'), ~ replace_na(.x, 0))) %>% 
  mutate(gender = as_factor(gender)) %>%
  mutate(gender = fct_relevel(gender, 'Male')) %>% 
  mutate(race = as_factor(race)) %>% 
  mutate(race = fct_relevel(race, 'White')) %>% 
  mutate(age = round(age / 10, 0) * 10) %>% 
  relocate(case_or_control, .after = last_col())
```

```{r message=FALSE}
library(caret)
require(doSNOW)
require(pROC)
```

```{r}
set.seed(7531)

# how to balance classes
class_balance_method <- c('downsample' = TRUE, 'upsample' = FALSE, 'asis' = FALSE)
stopifnot(sum(class_balance_method) == 1)
if (class_balance_method['downsample']){
  Xy <- downSample(Xy %>% select(-case_or_control), factor(Xy$case_or_control), yname='case_or_control')
} else if (class_balance_method['upsample']){
  Xy <- upSample(Xy %>% select(-case_or_control), factor(Xy$case_or_control), yname='case_or_control')
} else {
  # do nothing
}
```

```{r}
trainIndex <- createDataPartition(Xy$case_or_control, p = 0.80, list = FALSE)

df_ml_train <- Xy[trainIndex, ]
df_ml_test <- Xy[-trainIndex, ]
```

```{r}
# train a random forest, start with default
control <- trainControl(method='repeatedcv', 
                        number=5, 
                        repeats=3,
                        classProbs = TRUE)
mtry <- sqrt(ncol(df_ml_train))
tunegrid <- expand.grid(.mtry=mtry)

c1 <- makeCluster(6, type = "SOCK")
registerDoSNOW(c1)
rf_default <- train(case_or_control ~ ., 
                    data = df_ml_train %>% select(-medical_record_number) %>% mutate(case_or_control = factor(case_or_control)),
                    method = 'rf', 
                    metric = 'Kappa', 
                    truneGrid = tunegrid,
                    trControl = control)
stopCluster(c1)

importance <- varImp(rf_default)

feature_rank_rf <- as.data.frame(importance$importance) %>% rownames_to_column(var = 'feature') %>% arrange(-Overall)

prediction <- predict(rf_default, df_ml_test %>% select(-medical_record_number, -case_or_control))

confusionMatrix(prediction, factor(df_ml_test$case_or_control), positive = 'case')
```

```{r}
# use a boosting algorithm
c1 <- makeCluster(6, type = "SOCK")
registerDoSNOW(c1)
xgbGrid <- expand.grid(nrounds = c(30, 60, 90),
                       max_depth = c(10, 15, 20, 25),
                       colsample_bytree = seq(0.5, 0.9, length.out = 5),
                       eta = c(0.1, 0.2, 0.4),
                       gamma=0,
                       min_child_weight = 1,
                       subsample = c(0.5, 0.75)
                      )
xgboost_model <- train(case_or_control ~ ., 
                    data = df_ml_train %>% select(-medical_record_number) %>% mutate(case_or_control = factor(case_or_control)),
                    method = 'xgbTree', 
                    tuneGrid = xgbGrid,
                    trControl = control)
stopCluster(c1)
prediction <- predict(xgboost_model, df_ml_test %>% select(-medical_record_number, -case_or_control))

confusionMatrix(prediction, factor(df_ml_test$case_or_control), positive = 'case')
```