### 2022-04-22 code/r/02_E_numeric_vs_hpo_combined.Rmd

---
title: "numeric vs HPO"
output: html_notebook
---

In this page, we combine the analysis of numeric value vs HPO terms for prediction tasks.

```{r}
library(tidyverse)
library(glue)
```

Load previous data analysis results
```{r}
liver_analysis <- read.csv('cache/liver_analysis.csv')
crc_analysis <- read.csv('cache/crc_case_matched_controls_analysis.csv')
acute_kidney_failure_analysis <- read.csv('cache/acute_kidney_failure_analysis.csv')
anemia_analysis <- read.csv('cache/anemia_analysis.csv')
```

```{r}
numeric_vs_hpo <- bind_rows(liver_analysis %>% mutate(disease = 'abnormal liver function'),
                            crc_analysis %>% mutate(disease = 'colon cancer'), 
                            acute_kidney_failure_analysis %>% mutate(disease = 'acute kidney failure'), 
                            anemia_analysis %>% mutate(disease = 'aplastic anemia'))

numeric_vs_hpo %>% sample_n(4)

numeric_vs_hpo_final <- numeric_vs_hpo %>% 
  select(disease, loinc, estimate1, estimate2, p.value, interpretation, hpotermid) %>% 
  rename(auc_continuous = estimate1, auc_hpo = estimate2) %>% 
  group_by(disease, loinc) %>% 
  arrange(desc(auc_hpo)) %>% 
  mutate(seq = row_number()) %>% 
  ungroup() %>% 
  arrange(disease, loinc, seq) %>% 
  filter(seq == 1) %>% mutate(frac = auc_hpo/auc_continuous) %>% 
  mutate(frac_rm_rand = (auc_hpo - 0.5) / (auc_continuous - 0.5))

numeric_vs_hpo_final
```
```{r}
numeric_vs_hpo_final  %>% 
  #filter(estimate1 > 0.6) %>%
  summarise(frac_median = median(frac), frac_rm_rand_median = median(frac_rm_rand))
```
```{r}
df <- numeric_vs_hpo_final %>% 
  select(disease, loinc, auc_continuous, auc_hpo) %>% 
  pivot_longer(cols = starts_with('auc'), names_to = 'method', names_prefix = 'auc_', values_to = 'auc')

plot_AUC_differences_numeric_vs_hpo <- ggplot(df, aes(x = method, y = auc, group = interaction(disease, loinc))) + 
  geom_point(size = 1) +
  geom_line(aes(color = disease)) +
  scale_x_discrete(breaks = c('continuous', 'hpo'), labels = c('original numeric value', 'transformed to HPO')) +
  xlab("Use Lab Test Results As") +
  ylab("AUC") +
  theme_bw() + 
    theme(panel.grid = element_blank(), legend.position = 'none',
          axis.text = element_text(size = 8),
          axis.title = element_text(size = 8), 
          strip.text.x = element_text(margin = margin(0.3, 0, 0.3, 0), size = 5),
          plot.title = element_text(size = 8)) +
  facet_wrap(~ disease)

ggsave(filename = 'cache/plot_AUC_differences_numeric_vs_hpo.pdf', width = 6, height = 6)

plot_AUC_differences_numeric_vs_hpo
```