### 2022-04-22 code/r/03_biomarker.Rmd

---
title: "Biomarker screening"
output: html_notebook
---

In this section, we try to screen the phenotypes for biomarkers associated with pregression to severe asthma.

```{r message=FALSE}
library(tidyverse)
library(glue)
library(readxl)
require(DBI)
require(RPostgres)
require(egg)
require(gridExtra)
require(lubridate)
require(furrr)
require(dbplyr)

dsca <- DBI::dbConnect(RPostgres::Postgres(), 
                       host=Sys.getenv('DSCA_PHI_HOST'), 
                       dbname = Sys.getenv('DSCA_PHI_DBNAME'), 
                       port = Sys.getenv('DSCA_PHI_PORT'), 
                       user=Sys.getenv('DSCA_PHI_USERNAME'),
                       password= Sys.getenv('DSCA_PHI_PASSWORD'))

# cache directory to read from or write to
CACHE_DIR = '../../cache'

if (!dir.exists(CACHE_DIR)){
  dir.create(CACHE_DIR)
}
FORCE_RERUN <- TRUE
```

## Cohort 
The starting cohort are patients with any asthma (J45) and asthma-related diagnosis.

Within this starting cohort, we find one subset with severe asthma diagnosis. We further find a subset that has non-severe asthma diagnosis before they progressed to the severe form. These patients are the cases.

Another subcohort are the ones that have never been associated with severe asthma. These patients are the controls.

Final table has three columns:

- medical_record_number
- cohort: case or control
- t0: earliest severe asthma dx date for cases and last asthma diagnosis for controls

```{sql connection=dsca}
-- dx of asthma and asthma-related symptoms
create temp table asthma_related_dx as 
select medical_record_number , calendar_date , context_name , context_diagnosis_code 
from hai_az_prod.diagnosis_2020july dj 
where context_diagnosis_code ~* '^J45' or context_diagnosis_code ~* '^R06';
```

```{sql connection=dsca}
-- severe asthma
create temp table severe_asthma_patients as
select medical_record_number , min(calendar_date) as min_calendar_date
from hai_az_prod.diagnosis_2020july dj 
where context_diagnosis_code ~* '^J45\\.5'
group by medical_record_number;  
```

A subset of severe asthma patients who were progressed from non-severe to severe. These are cases. 
```{sql connection=dsca}
-- severe asthma progressed from non-severe: 2593
create temp table progressed_to_severe_asthma_patients as
select distinct medical_record_number 
from hai_az_prod.diagnosis_2020july dj 
join severe_asthma_patients using (medical_record_number)
where (context_diagnosis_code ~* '^J45(\\.)?[2349]' or context_diagnosis_code ~* '^R06') and calendar_date < min_calendar_date;
```

Patients who had asthma but not severe. These are controls. 
```{sql connection=dsca}
-- control : 323420
create temp table never_severe_asthma_patients as
select medical_record_number, max(calendar_date) as max_calendar_date, random() as random 
from asthma_related_dx
where medical_record_number not in (select medical_record_number from severe_asthma_patients)
group by medical_record_number ;
```

Combine cases and controls in one table. For now, we will use the 
```{sql connection=dsca}
-- combined cohort 
create temp table cohort as 
select medical_record_number , min_calendar_date as t0, 'case' as case_or_control
from severe_asthma_patients 
join progressed_to_severe_asthma_patients using (medical_record_number)
union all
select medical_record_number , max_calendar_date as t0, 'control' as case_or_control
from never_severe_asthma_patients
where random < 2593.0 * 2 / 323420;
```

## Extraction of data

We need to extract the following information:

- the cohort table: medical_record_number, cohort, t0
- tested phenotypess: medical_record_number, termid, testdate
- observed phenotypes: medical_record_number, termid, testdate
- demographics: medical_record_number, sex, race, birthdate

Get data from the database. The following method are lazily executed.
```{r}
cohort <- tbl(dsca, 'cohort')
tested_phenotypes <- tbl(dsca, sql("
    with testable_phenotypes_all as (
    	select lab.medical_record_number , lab.context_name , lab.lab_date , lab.lab_time , lab.loinc , testable.termid , testable.is_tested_directly
    	from hai_az_prod.scc_and_epic_lab_after_loinc2hpo_2020july lab
    	join hai_az_prod.loinc_testable_phenotypes testable on lab.loinc = testable.loincid
    	join cohort using (medical_record_number)
      )
     select medical_record_number, lab_date, termid
     from testable_phenotypes_all"))
observed_phenotypes <- tbl(dsca, sql("
    with abnormal_lab as (
      select lab.*
      from hai_az_prod.scc_and_epic_lab_after_loinc2hpo_2020july lab
      join cohort using (medical_record_number)
      where hpotermid is not null and not isnegated),
    abnormal_lab_after_infer as (
      SELECT l.medical_record_number, l.lab_date, date_part(year, lab_date) as lab_year, l.local_test_code, l.loinc, l.hpotermid, l.isnegated, pair.ancestor , pair.distance
      FROM abnormal_lab l
      left join hai_az_prod.hpo_is_a_pairs pair on l.hpotermid = pair.current)
    select medical_record_number , lab_date, ancestor as termid
    from abnormal_lab_after_infer"))
cohort_demographics <- tbl(dsca, sql("
    select person.*
    from hai_az_prod.person_msdw_2020july person
    join cohort using (medical_record_number)"))

```

## Data transformation

- tested phenotypes: calculate testdate relative to t0; cut the relative date into yearly brackets and aggregated on yearly basis
- observed phenotypes: similar to the above
- demographics: calculate age at the time of t0

```{r}
cohort_tested_phenotypes_yearly_count <- tested_phenotypes %>%
  left_join(cohort, by = 'medical_record_number') %>%
  mutate(lab_date_rel = interval(t0, lab_date) / ddays()) %>% 
  filter(lab_date_rel < 0) %>%
  filter(lab_date_rel > -365 * 3) %>%
  mutate(lab_date_rel_bracket = cut(lab_date_rel / 365, breaks=seq(-3,0))) %>%
  group_by(medical_record_number, lab_date_rel_bracket, termid) %>% 
  summarise(termid_count = n(), .groups = 'drop')
```
```{r}
cohort_observed_phenotypes_yearly_count <- observed_phenotypes %>% 
  left_join(cohort, by = 'medical_record_number') %>%
  mutate(lab_date_rel = interval(t0, lab_date) / ddays()) %>%
  filter(lab_date_rel < 0) %>%
  filter(lab_date_rel > -365 * 3) %>%
  mutate(lab_date_rel_bracket = cut(lab_date_rel / 365, breaks=seq(-3,0))) %>%
  group_by(medical_record_number, lab_date_rel_bracket, termid) %>% 
  summarise(termid_count = n(), .groups = 'drop')
```

```{r}
## collect all the phenotypes in one dataframe
phenotype_features <- cohort_tested_phenotypes_yearly_count %>% 
  rename(tested_count = termid_count) %>%
  left_join(cohort_observed_phenotypes_yearly_count %>% rename(observed_count = termid_count), 
            by = c('medical_record_number', 'lab_date_rel_bracket', 'termid')) %>%
  mutate(observed_count = ifelse(is.na(observed_count), 0, observed_count)) %>% 
  mutate(is_tested = tested_count > 0) %>%
  mutate(is_observed = observed_count > 0)
```

The final data frame should have the following columns:
medical_record_number: patient id 
termid: phenotype
is_tested: TRUE or FALSE
is_observed: TRUE or FALSE
time_window: [-3, -2), [-2, -1), [-1, 0)
sex
race
age: age at first severe asthma diagnosis (cases) or last asthma diagnosis (control)
outcome: progressed to severe asthma or not

## Statistical analysis
For the initial analysis, the model is as below:

outcome ~ is_tested + is_tested:is_observed

```{r}
confounders_outcome <- cohort %>% 
  inner_join(cohort_demographics, by = 'medical_record_number') %>%
  mutate(age = interval(date_of_birth, t0) / dyears()) %>%
  select(medical_record_number, gender, race, age, case_or_control) %>% 
  filter(!is.na(age))

d <- phenotype_features %>% 
  filter(termid == 'HP:0001880') %>%
  filter(lab_date_rel_bracket == '(-1,0]')

d <- phenotype_features %>% 
  filter(termid == 'HP:0001873') %>%
  filter(lab_date_rel_bracket == '(-1,0]')

prepare_Xy <- function(df_phenotype_feature, df_confounders_outcome){
  df_model <- df_confounders_outcome %>% 
    left_join(df_phenotype_feature, by = 'medical_record_number') %>%
    select(medical_record_number, gender, race, age, tested_count, observed_count, is_tested, is_observed, case_or_control) %>%
    ## fill in missing values
    replace_na(list(tested_count = 0, observed_count = 0, is_tested = FALSE, is_observed = FALSE)) %>%
    ## factorize categorical variables
    mutate(gender = factor(gender)) %>%
    mutate(race = factor(race)) %>%
    mutate(is_tested = factor(is_tested)) %>%
    mutate(is_observed = factor(is_observed)) %>%
    mutate(case_or_control = factor(case_or_control)) %>%
    ## set the reference levels for categorical variables
    mutate(gender = relevel(gender, ref='Male')) %>%
    mutate(race = relevel(race, ref='White')) %>%
    mutate(is_tested = relevel(is_tested, ref = 'FALSE')) %>%
    mutate(is_observed = relevel(is_observed, ref = 'FALSE')) %>%
    mutate(case_or_control = relevel(case_or_control, ref = 'control'))
  return (df_model)
}

glm_Xy <- function(Xy){
  model <- glm(formula = 'case_or_control ~ gender + race + age + is_tested / is_observed', data = Xy, family = binomial(link = 'logit'))
  model_coeff <- broom::tidy(model)
  return (model_coeff)
}

df_model <- prepare_Xy(df_phenotype_feature = d %>% select(-lab_date_rel_bracket, -termid), 
                       df_confounders_outcome = confounders_outcome)

# df_model %>% sample_n(10)

glm_Xy(df_model) %>% filter(str_starts(term, 'is_testedTRUE'))

fit_model <- function(df_phenotype_feature, df_confounders_outcome) {
  df_model <- prepare_Xy(df_phenotype_feature, df_confounders_outcome)
  model_coeff <- data.frame(term = character(), estimate = double(), std.error = double(), statistic = double(), p.value = double())
  tryCatch({model_coeff = glm_Xy(df_model)}, error = function(e) warning('cannot fit model'))
  #error = function(e) print('model did not run'))
  return (model_coeff)
}

fit_model(df_phenotype_feature = d %>% select(-lab_date_rel_bracket, -termid), 
                       df_confounders_outcome = confounders_outcome)

## explore other statistical models
# model2 <- glmer(formula = 'case_or_control ~ gender + race + age + is_observed + (1 + is_observed | is_tested)', data = df_model, family = binomial(link = 'logit') )
# summary(model2)
```

Screen all the phenotypes

```{r}
tic <- Sys.time()

plan(multisession, workers = 8)

screen_results <- phenotype_features %>% 
  #filter(termid == 'HP:0001880') %>%
  #filter(lab_date_rel_bracket == '(-1,0]') %>%
  group_by(lab_date_rel_bracket, termid) %>%
  nest() %>%
  #calling ungroup() is critical for furrr to work in parallel
  ungroup() %>%
  mutate(coefficient = future_map(data, fit_model, df_confounders_outcome = confounders_outcome)) %>%
  select(-data) %>%
  unnest(coefficient)

toc <- Sys.time()

toc - tic
```
```{r}
## how many biomarkers we have screened
## at each time window
N_BIOMARKERS_SCREENED <- screen_results %>% select(lab_date_rel_bracket, termid) %>% distinct() %>% nrow()
screen_results %>% group_by(lab_date_rel_bracket) %>% summarise(n = n_distinct(termid))
```

Annotate the screen results by adding the HPO term label, and also create a few different views that rank the results differently. 
```{r}
hpo_term_list <- read.csv('../../in/hp_term_list.csv', header = FALSE, col.names = c('termid', 'distance_to_root', 'label'), stringsAsFactors = FALSE)

screen_results_annotated <- screen_results %>%  filter(str_starts(term, 'is_testedTRUE')) %>% arrange(lab_date_rel_bracket) %>%
  left_join(hpo_term_list, by = 'termid') %>%
  select(lab_date_rel_bracket, termid, label, distance_to_root, term, distance_to_root, estimate, std.error, statistic, p.value)

write.csv(screen_results_annotated, file = glue('{CACHE_DIR}/screen_results_annotated.csv'), row.names = FALSE)

screen_results_annotated %>% head()
```

```{r}
screen_results_annotated_formatted <- screen_results_annotated %>% 
  mutate(OR = exp(estimate)) %>% 
  mutate(OR = round(OR, 2)) %>% 
  #mutate(p.value = scales::scientific(p.value, digits = 3)) %>% 
  select(-estimate, -std.error, -statistic) %>% 
  mutate(term = case_when(term == 'is_testedTRUE' ~ 'tested',
                          term == 'is_testedTRUE:is_observedTRUE' ~ 'observed',
                          TRUE ~ 'NEVER GONNA HAPPEN')) %>% 
  pivot_wider(id_cols = c(lab_date_rel_bracket, termid, label, distance_to_root), names_from = term, values_from = c(OR, p.value)) %>% 
  relocate(p.value_tested, .after = 'OR_tested') %>% 
  rename(`Odds Ratio (Been Tested)` = OR_tested) %>% 
  rename(`p value (Been Tested)` = p.value_tested) %>% 
  rename(`Odds Ratio (Been Observed)` = 'OR_observed') %>% 
  rename(`p value (Been Observed)` = 'p.value_observed')

write.csv(screen_results_annotated_formatted, file = glue('{CACHE_DIR}/screen_results_annotated.csv', row.names = FALSE)

screen_results_annotated_formatted %>% head()
```

Highlight some biomarkers:

Eosinophilia
```{r}
eosinophilia <- screen_results_annotated_formatted %>% 
  filter(label == 'Eosinophilia') %>% 
  select(-distance_to_root) %>% 
  mutate(across(starts_with('p value'), scales::scientific))

sjPlot::tab_df(eosinophilia %>% rename(`Lab Year Bracket` = lab_date_rel_bracket), 
                title = 'Table Association of Eosinophilia with Severe Asthma',
               file = glue('{CACHE_DIR}/table_eosinophilia_with_severe_asthma.html',
               alternate.rows = TRUE)

eosinophilia
```

Rank by Odds Ratio (Been Observed)
```{r}
significant_by_tested_and_observed <- screen_results_annotated_formatted %>% 
  arrange(desc(`Odds Ratio (Been Observed)`)) %>% 
  filter(`p value (Been Tested)` < 0.05/N_BIOMARKERS_SCREENED) %>%
  filter(`p value (Been Observed)` < 0.05/N_BIOMARKERS_SCREENED) %>%
  filter(`Odds Ratio (Been Tested)` > 1) %>% 
  filter(`Odds Ratio (Been Observed)` > 1) %>% 
  filter(lab_date_rel_bracket == '(-1,0]') %>%
  mutate(across(starts_with('p value'), scales::scientific))

significant_by_tested_and_observed_deduped <- significant_by_tested_and_observed %>% 
  rows_delete(tibble(label = c('Abnormal eosinophil morphology', 
                               'Abnormal leukocyte morphology',
                               'Abnormal immune system morphology',
                               'Abnormal cellular immune system morphology',
                               'Abnormal leukocyte count',
                               'Abnormal granulocyte morphology',
                               'Abnormality of the respiratory system',
                               'Functional respiratory abnormality')), by = 'label')

significant_by_tested_and_observed_deduped
```
```{r}
significant_by_tested_and_observed_deduped_three_years <- screen_results_annotated_formatted %>% 
  filter(termid %in% significant_by_tested_and_observed_deduped$termid) %>% 
  mutate(termid = factor(termid, levels = significant_by_tested_and_observed_deduped$termid)) %>% 
  arrange(termid, lab_date_rel_bracket) %>%
  mutate(across(starts_with('p value'), scales::scientific))

significant_by_tested_and_observed_deduped_three_years
```

```{r}
sjPlot::tab_df(significant_by_tested_and_observed_deduped %>% select(-distance_to_root, -lab_date_rel_bracket), 
                title = 'Table Selected Phenotypes (within 1 year prior) Associated with Severe Asthma',
               file = glue('{CACHE_DIR}/table_selected_biomarker_from_labs_within_1_year_prior_to_dx.html', 
               alternate.rows = TRUE)
sjPlot::tab_df(significant_by_tested_and_observed_deduped_three_years %>% 
                 select(-distance_to_root) %>% rename(`Lab Year Bracket` = lab_date_rel_bracket), 
                title = 'Table Selected Phenotypes (within past 3 years) Associated with Severe Asthma',
               file = glue('{CACHE_DIR}/table_selected_biomarker_from_labs_within_3_years_prior_to_dx.html',
               alternate.rows = TRUE)
```

## Cache data for future use
```{r}
cache_data <- TRUE

if (cache_data){
  cache_list <- list(confounders_outcome = confounders_outcome,
                     phenotype_features = phenotype_features)

  for (i in 1:length(cache_list)){
    filename <- names(cache_list)[[i]]
    write.table(cache_list[[i]], file = glue('{CACHE_DIR}/{filename}.csv'), sep = ',', row.names = FALSE, col.names = TRUE)
  }
}
```