### 2022-04-27 code/r/02_E_numeric_vs_hpo_combined.Rmd

---
title: "numeric vs HPO"
output:
  html_document:
    df_print: paged
---

In this page, we combine the analysis of numeric value vs HPO terms for prediction tasks.

```{r}
library(tidyverse)
library(glue)

source('functions.R')

# cache directory to read from or write to
CACHE_DIR = '../../cache'

# output directory to write analysis results
OUT_DIR = '../../out'
if (!dir.exists(OUT_DIR)){
  dir.create(OUT_DIR)
}

```

Load previous data analysis results
```{r}
liver_analysis <- read.csv(glue('{OUT_DIR}/liver_analysis.csv'))
crc_analysis <- read.csv(glue('{OUT_DIR}/crc_case_matched_controls_analysis.csv'))
acute_kidney_failure_analysis <- read.csv(glue('{OUT_DIR}/acute_kidney_failure_analysis.csv'))
anemia_analysis <- read.csv(glue('{OUT_DIR}/anemia_analysis.csv'))
```

```{r}
numeric_vs_hpo <- bind_rows(liver_analysis %>% mutate(disease = 'abnormal liver function'),
                            crc_analysis %>% mutate(disease = 'colon cancer'), 
                            acute_kidney_failure_analysis %>% mutate(disease = 'acute kidney failure'), 
                            anemia_analysis %>% mutate(disease = 'aplastic anemia'))

numeric_vs_hpo %>% sample_n(4)

numeric_vs_hpo_final <- numeric_vs_hpo %>% 
  select(disease, loinc, estimate1, estimate2, p.value, interpretation, hpotermid) %>% 
  rename(auc_continuous = estimate1, auc_hpo = estimate2) %>% 
  group_by(disease, loinc) %>% 
  arrange(desc(auc_hpo)) %>% 
  mutate(seq = row_number()) %>% 
  ungroup() %>% 
  arrange(disease, loinc, seq) %>% 
  filter(seq == 1) %>% mutate(frac = auc_hpo/auc_continuous) %>% 
  mutate(frac_rm_rand = (auc_hpo - 0.5) / (auc_continuous - 0.5))

numeric_vs_hpo_final
```
```{r}
numeric_vs_hpo_final  %>% 
  #filter(estimate1 > 0.6) %>%
  summarise(frac_median = median(frac), frac_rm_rand_median = median(frac_rm_rand))
```
```{r}
df <- numeric_vs_hpo_final %>% 
  select(disease, loinc, auc_continuous, auc_hpo) %>% 
  pivot_longer(cols = starts_with('auc'), names_to = 'method', names_prefix = 'auc_', values_to = 'auc')

plot_AUC_differences_numeric_vs_hpo <- ggplot(df, aes(x = method, y = auc, group = interaction(disease, loinc))) + 
  geom_point(size = 1) +
  geom_line(aes(color = disease)) +
  scale_x_discrete(breaks = c('continuous', 'hpo'), labels = c('numeric', 'HPO')) +
  xlab("Use Lab Test Results As") +
  ylab("AUC") +
  theme_bw() + 
    theme(panel.grid = element_blank(), legend.position = 'none',
          axis.text = element_text(size = 8),
          axis.title = element_text(size = 8), 
          strip.text.x = element_text(margin = margin(0.3, 0, 0.3, 0), size = 8),
          plot.title = element_text(size = 8)) +
  facet_wrap(~ disease)

ggsave2(filenames = c(glue('{OUT_DIR}/plot_AUC_differences_numeric_vs_hpo.png'), glue('{OUT_DIR}/plot_AUC_differences_numeric_vs_hpo.pdf')), plot = plot_AUC_differences_numeric_vs_hpo, width = 3.2, height = 3.2)

plot_AUC_differences_numeric_vs_hpo
```