### 2022-04-28 code/r/01_B.Rmd

---
title: "01_B_observed_normalized_by_tested"
output: html_document
date: '2022-04-27'
author: Aaron Zhang
---

This is a small extension for the main analysis of "01_MSDW_HPO_Phenotype_Analysis.Rmd". 
The goal is to analyze the trendlines of observed phenotypes but now normalize by how many patients were tested for each corresponding phenotype.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(glue)
source('functions.R')
```

## load data
```{r}
# cache directory to read from or write to
CACHE_DIR = '../../cache'
if (!dir.exists(CACHE_DIR)){
  dir.create(CACHE_DIR)
}

# output directory to write analysis results
OUT_DIR = '../../out'
if (!dir.exists(OUT_DIR)){
  dir.create(OUT_DIR)
}

tested_phenotype_rank_all_by_year <- read_csv(glue('{CACHE_DIR}/tested_phenotype_rank_all_by_year.csv'))
observed_phenotype_rank_all_by_year <- read_csv(glue('{CACHE_DIR}/observed_phenotype_rank_all_by_year.csv'))
```

## analysis
Calculate fraction of patients observed for each phenotype by patients tested for that phenotype. 
```{r}
observed_divide_tested_all_by_year <- observed_phenotype_rank_all_by_year %>% 
  rename(n_patient_observed = n_patient) %>% 
  select(-n_record) %>%
  left_join(tested_phenotype_rank_all_by_year %>% 
              rename(n_patient_tested = n_patient) %>% 
              select(termid, lab_year, n_patient_tested),
            by = c('termid', 'lab_year')) %>%
  relocate(n_patient_tested, .after = n_patient_observed) %>%
  mutate(frac_observed_tested = n_patient_observed / n_patient_tested)
```

Below, we plot level HPO terms.
```{r}
plot_phenotypes_of_interest_by_year <- function(terms_of_interest, df){
  hpo_terms_selected <- df %>% filter(termid %in% terms_of_interest) %>% select(termid, label) %>% distinct()

  y_str = 'frac_observed_tested'
  y_label = 'patient fraction (observed / tested)'

  p <- df %>% filter(termid %in% terms_of_interest) %>% 
    filter(lab_year <= 2019 & lab_year > 2000) %>%
    ggplot(aes(x = lab_year, y = eval(parse(text=y_str)), group = termid, color = termid)) + 
    geom_point() +
    geom_line() +
    scale_color_discrete(breaks = hpo_terms_selected$termid, labels = str_c(hpo_terms_selected$termid, hpo_terms_selected$label, sep = " ")) +
    xlab('Year') +
    ylab(y_label) +
    theme_az()

  return (p)
}
```

```{r}
hpo_level_1_terms <- observed_phenotype_rank_all_by_year %>% filter(distancetoroot == 1) %>% select(termid, label) %>% distinct()

p <- plot_phenotypes_of_interest_by_year(terms_of_interest = hpo_level_1_terms$termid, df=observed_divide_tested_all_by_year)

p
```

Run statistical testing of the trendlines. It is probably a better metric to claim whether the cohort is more or less healthier for one phenotype as this is normalized by patients tested for that phenotype.

```{r output.max=-1}
trendline_lm_test <- observed_divide_tested_all_by_year %>% 
  filter(distancetoroot == 1) %>% 
  filter(lab_year > 2000) %>%
  rename(fraction = frac_observed_tested) %>%
  select(termid, lab_year, label, fraction) %>% 
  group_by(termid, label) %>% 
  nest() %>% 
  mutate(lm_model = map(data, function(d) lm(fraction ~ lab_year, data = d))) %>%
  mutate(lm_coeff = map(lm_model, broom::tidy)) %>%
  select(-data, -lm_model) %>% 
  unnest(lm_coeff) %>% 
  filter(term == 'lab_year') %>% 
  mutate(p.value.bonferroni = p.value * 11)

trendline_lm_test_formatted <- trendline_lm_test %>% select(termid, label, estimate, p.value, p.value.bonferroni) %>% 
  arrange(p.value.bonferroni) %>%
  rename(beta = estimate) %>% 
  mutate(significance = symnum(p.value.bonferroni, cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, Inf), symbols = c("***", "**", "*", ".", " ")) ) %>% 
  mutate(across(.cols = function(x) is.double(x), function(x) scales::scientific(x, digits = 3)))

write.csv(trendline_lm_test_formatted, file = glue('{OUT_DIR}/table_observed_normalized_by_tested_phenotypes_by_year_level_1_normalized_trendline_lm_model.csv'), row.names=FALSE)

trendline_lm_test_formatted
```

In the last section, we show two sets of phenotypes. Set 1, those phenotype trend lower when looking at observed phenotypes only; Set 2, those phenotype trend higher when looking at observed phenotypes only.

The terms have significantly decreased trendlines when looking at patient counts of observed phenotypes. 
```{r}
hp_terms_of_interest <- c('HP:0001871','HP:0002715','HP:0025031','HP:0001626')
p <- plot_phenotypes_of_interest_by_year(terms_of_interest = hp_terms_of_interest, df=observed_divide_tested_all_by_year)

ggsave2(filenames = c(glue('{OUT_DIR}/plot_observed_by_tested_set1.png'), glue('{OUT_DIR}/plot_observed_by_tested_set1.pdf')), plot = p, width = 8, height = 4)

p
```

The terms have significantly increased trendlines when looking at patient counts of observed phenotypes. 
```{r}
hp_terms_of_interest <- c('HP:0000119','HP:0001939')
p <- plot_phenotypes_of_interest_by_year(terms_of_interest = hp_terms_of_interest, df=observed_divide_tested_all_by_year)

ggsave2(filenames = c(glue('{OUT_DIR}/plot_observed_by_tested_set2.png'), glue('{OUT_DIR}/plot_observed_by_tested_set2.pdf')), plot = p, width = 8, height = 4)

p
```