### 2022-04-28 code/r/01_C.Rmd

---
title: "01_B_racial_disparity_observed_normalized_by_tested"
output: html_document
date: '2022-04-28'
author: Aaron Zhang
---

This is another small extension for the main analysis of "01_MSDW_HPO_Phenotype_Analysis.Rmd". 
The goal is to analyze the trendlines of observed phenotypes for each race but now normalize by how many patients were tested for each corresponding phenotype.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(glue)
source('functions.R')
```

## load data
```{r}
# cache directory to read from or write to
CACHE_DIR = '../../cache'
if (!dir.exists(CACHE_DIR)){
  dir.create(CACHE_DIR)
}

# output directory to write analysis results
OUT_DIR = '../../out'
if (!dir.exists(OUT_DIR)){
  dir.create(OUT_DIR)
}

IN_DIR = '../../in'

observed_phenotype_rank_all_by_year_race_sex <- read_csv(glue('{CACHE_DIR}/observed_phenotype_rank_all_by_year_race_sex.csv'))
tested_phenotype_rank_all_by_race_year_sex <- read_csv(glue('{CACHE_DIR}/tested_phenotype_rank_all_by_race_year_sex.csv'))
hp_term_list <- read.csv(glue('{IN_DIR}/hp_term_list.csv'), header = FALSE, col.names = c('termid', 'distancetoroot', 'label'))
```

## analysis
Calculate fraction of patients observed for each phenotype by patients tested for that phenotype.

```{r}
observed_phenotype_rank_all_by_year_race <- observed_phenotype_rank_all_by_year_race_sex %>% 
  group_by(termid, lab_year, race) %>%
  summarise(n_patient_observed = sum(n_patient), .groups = 'drop')

tested_phenotype_rank_all_by_race_year <- tested_phenotype_rank_all_by_race_year_sex %>%
  group_by(termid, lab_year, race) %>%
  summarise(n_patient_tested = sum(n_patient), .groups = 'drop')
```

```{r}
observed_normalized_by_tested_all_year_race <- observed_phenotype_rank_all_by_year_race %>% 
  left_join(tested_phenotype_rank_all_by_race_year, by = c('termid', 'lab_year', 'race')) %>%
  mutate(frac_observed_tested = n_patient_observed / n_patient_tested)

observed_normalized_by_tested_all_year_race <- observed_normalized_by_tested_all_year_race %>% 
  left_join(hp_term_list, by = 'termid')
```

Below, we plot level HPO terms.
```{r}
plot_phenotypes_of_interest_by_year <- function(terms_of_interest, df){
  hpo_terms_selected <- df %>% filter(termid %in% terms_of_interest) %>% select(termid, label) %>% distinct()

  y_str = 'frac_observed_tested'
  y_label = 'patient fraction (observed / tested)'

  df <- df %>% mutate(termid_label = str_c(label, glue('[{termid}]'), sep = '\n'))

  p <- df %>% filter(termid %in% terms_of_interest) %>% 
    filter(lab_year <= 2019 & lab_year > 2000) %>%
    #ggplot(aes(x = lab_year, y = eval(parse(text=y_str)), group = termid, color = termid)) + 
    ggplot(aes(x = lab_year, y = eval(parse(text=y_str)), group = race, color = race)) +
    geom_point() +
    geom_line() +
    #scale_color_discrete(breaks = hpo_terms_selected$termid, labels = str_c(hpo_terms_selected$termid, hpo_terms_selected$label, sep = " ")) +
    xlab('Year') +
    ylab(y_label) +
    theme_az() + 
    facet_wrap(~termid_label, ncol = 2)

  return (p)
}
```

```{r}
hpo_level_1_terms <- observed_normalized_by_tested_all_year_race %>% filter(distancetoroot == 1) %>% select(termid, label) %>% distinct()

p <- plot_phenotypes_of_interest_by_year(terms_of_interest = hpo_level_1_terms$termid, df=observed_normalized_by_tested_all_year_race %>% filter(race %in% c('White', 'Black or African American', 'Asian')))

ggsave2(filenames = c(glue('{OUT_DIR}/plot_racial_difference_observed_normalized_by_tested.png'), glue('{OUT_DIR}/plot_racial_difference_observed_normalized_by_tested.pdf')), width=8, height = 10)
```

The plot shows racial differences of observed phenotypes.

![racial differences](./../../out/plot_racial_difference_observed_normalized_by_tested.png)